Forecasting Gamma-Ray Bursts using Gravitational Waves

### 2018-04-15 Paper/Manuscript.tex
Corrections, new references. Added one paper as a PDF.

\begin{abstract}
We explore the intriguing possibility of employing future ground-based gravitational-wave interferometers to detect the inspiral of binary neutron stars early enough to alert electromagnetic observatories so that a gamma-ray burst (GRB) can be observed in its entirety from its very start.
We quantify the ability to predict a GRB by computing the time a binary neutron star (BNS) system takes to inspiral from its moment of detection to its final merger. We define the moment of detection to be the instant at which the inteferometer network accumulates a signal-to-noise ratio of 15. %for the BNS inspiral.
For our computations, %of advance warning times, 
we specifically consider BNS systems at luminosity distances of (i) $D\le200\,$Mpc in the three-interferometer Advanced-LIGO-Virgo network of 2020, and (ii) $D \le 1000\,$Mpc in the Einstein Telescope's B and C configurations. 
In the case of Advanced LIGO-Virgo we find that we may at best get a few minutes of warning time thus we expect no forecast of GRBs in the 2020s. 
On the other hand, Einstein Telescope will provide us with advance warning times up to five hours.  
Taking one hour as a benchmark warning time, we obtain a corresponding horizon
distance of roughly 600 Mpc for the Einstein Telescope C configuration.
%, more specifically, greater than an hour
%for BNS inspirals closer than roughly 600\,Mpc. Taking this luminosity distance to be our horizon distance, 
Using current BNS event rates within this horizon distance, we show that this translates to forecasts of $\gtrsim \ord(10^2)$ GRBs in the 2030s. %with Einstein C. %with the C configuration of Einstein Telescope.
We reapply our warning-time computation to binary black hole - neutron star inspirals and find that we expect 1 to 3 tidal disruption events 
to be forecast by the same detector.
%Our main conclusions are: (i) we expect ALV network to forecast 1 to 3 GRBs in the 2020s, (ii)

@@

GW170817 was initially identified by the LIGO-Hanford interferometer (H1) 
using a template bank of gravitational waveforms computed within the framework of post-Newtonian theory \cite{Buonanno:2009zt, Blanchet_LRR}.
The inspiral swept across the Advanced Ligo-Virgo (ALV) network's bandwidth from approximately $ 30\,$Hz to about $ 2\,$kHz in 57 seconds
executing nearly $ 3000$ gravitational wave cycles and accumulating a network signal-to-noise ratio of 32.4 with a false-alarm rate of one per $8.0\times 10^4$ years \cite{GW170817}. 
The total mass of the system was inferred to be $2.74^{+0.04}_{-0.01} M_\odot$ with component masses in the range $1.17-1.60 M_\odot$ consistent
with neutron stars (assuming low spin priors, see Sec.~\ref{sec:idealizations}).
Moreover, both the gravitational wave (GW) and EM observations were consistent with a source luminosity distance of $\sim 40\,$Mpc
and supported the hypothesis that GW170817 resulted from the inspiral and merger of two neutron stars, 
causing the prompt short-hard gamma-ray burst (GRB170817A) in NGC4993 as first suggested by Ref.~\cite{Eichler:1989ve}.

@@

Given the sensitivities of interferometers to date, efforts have thus far mostly focused on reducing the latency of EM follow-ups, 
i.e., the lag time between the merger and the alerting of the EM observatories. 
LIGO-Virgo science runs in 2009-2010 yielded a latency of $\sim 30$ to $60\,$minutes \cite{Abbott:2011ys} (no detections).
In the case of GW170817, the latency was 40\,minutes and $36^{+1.7}\,$seconds \cite{GBM:2017lvd}.
It then took an additional $ 10$ hours to locate the optical transient. %designated AT 2017gfo.

@@

This article is organized as follows. Sec.~\ref{sec:BNS_inspiral} introduces the formulation for the leading-order evolution of the binary inspiral.
Sec.~\ref{sec:IFO_response} details the interferometer response to GWs based on topology.
Sec.~\ref{sec:idealizations} lists the various idealizations we employ to simplify our treatment and supplies justification for each.
Sec.~\ref{sec:GW170817} tests our model network and evolution using the parameters of GW170817 and the corresponding observations.
Sec.~\ref{sec:results} contains our main results presented as advance warning times in Tables~\ref{table:LIGO2020}, \ref{table:ET}; 
and as horizon distances and event rates in Table~\ref{table:horizon}.
In Sec.~\ref{sec:BH_NS}, we recompute the advance warning times in the case of black hole - neutron star inspirals to see whether or not 
future detectors can forecast a tidal disruption event, for which we summarize our findings in Table~\ref{table:ET_BH_NS}.

@@

\section{Binary neutron star inspirals}\label{sec:BNS_inspiral}
\subsection{Newtonian evolution}
For us, the starting point for the evolution of a binary under gravitational radiation reaction
%the emission of GWs 
is the expression for the power emitted in GWs given by 
the celebrated Einstein quadrupole formula \cite{1918SPAW}
%for the generated by the time-variance of the mass quadrupole moment
%
\be
\dot{E} = \f{G}{5c^5}\langle \dddot{Q}_{ij}\dddot{Q}_{ij} \rangle \label{eq:Edot_quadrupole}\, .
\ee
%
Here, $Q_{ij}$ is a trace-reversed mass quadrupole moment which
is the dominant contribution to the power emitted %in gravitational radiation
by a system with changing mass moments.

Here, we are specifically interested in the motion of two point masses in a circular orbit around a common center of mass.
Labelling the masses by $m_1 \le m_2$, we have, for a binary in a circular orbit with Keplerian angular frequency $\Omega$ and separation $r$,
%
\be
\dot{E} = \f{32}{5}\f{G\mu^2}{c^5}\, r^4 \Omega^6\label{eq:Edot},
\ee
%
where $\mu= m_1 m_2/M$ is the reduced mass and $M=m_1+m_2$ is the total mass. 
We justify our simplification of motion to circular orbits in Sec.~\ref{sec:idealizations}.
%Our assumption of the circularity of the motion is well justified in

An important quantity in GW astronomy is the chirp mass of the binary given by
%
\be
M_c = \mu^{3/5} M^{2/5} = \f{(m_1 m_2)^{3/5}}{(m_1+m_2)^{1/5}} \label{eq:chirp_mass}\, ,
\ee
%
which turns Eq.~(\ref{eq:Edot}) into
%
\be
\dot{E} = \f{32}{5}\f{c^5}{G} \left(\f{G M_c\, \omega}{2c^3}\right)^{10/3}, \label{eq:Edot2}
\ee
%
where $\omega=2\Omega$ is the angular frequency of the quadrupolar GWs. 
The quantity in parantheses in Eq.~(\ref{eq:Edot2}) is ubiquitous in post-Newtonian theory and is
referred to as the dimensionless inverse separation $x$.
As $x\propto v^2/c^2$, it is used to keep track of post-Newtonian orders and we can see from in Eq.~(\ref{eq:Edot2}) that $\dot{E}\propto x^5$. 
%hence the common statement in the literature saying that gravitational radiation starts

@@

Let us now turn our attention to the GWs generated by the inspiral.
These are tensorial perturbations propagating at the speed of light in flat (Minkowski) spacetime.
In the so-called wave zone, defined by the condition $ c/f \ll D$, where $D$ is the distance to the GW source, 
these tensor modes satisfy the flat spacetime sourceless wave equation. %$\Box\bar{h}_{\mu\nu} = 0$.
%meaning that we are considering the solutions far from the GW source, 
In a suitable gauge, such as the commonly used transverse-traceless gauge, 
it can be shown that there exist only two physical degrees of freedom which manifest themselves as two independent polarization amplitudes $h_+$ and $h_\times$. %when interacting with masses away from the source.
For a binary in a circular orbit, these states read
%
\begin{align}
 h_+(t) &= h_c(t)\, \left(\f{1+\cos^2\iota}{2}\right)\, \cos[\Phi_\text{N}(t)],\label{eq:hplus_TD}\\
 h_\times(t) & =h_c(t)\,\cos\iota \sin[\Phi_\text{N}(t)]\label{eq:hcross_TD},
\end{align}
%
%
where
%
\be
h_c(t) =\f{4}{D}\left(\f{G M_c}{c^2}\right)^{5/3}\, \left(\f{\pi f(t)}{c}\right)^{2/3} \label{eq:h_c}
\ee
is the characteristic strain and
$\iota=\cos^{-1}(\mathbf{\hat{n}}\cdot{\mathbf{\hat{L}}}) $ is the inclination angle between the line of sight unit vector $\mathbf{\hat{n}} $ and the orbital angular momentum unit vector $\mathbf{\hat{L}}$. 
For an orbit seen edge on, $\iota = \pi/2$ giving $h_\times=0$, i.e,  a purely plus-polarized wave.
When $\iota=0$ the orbit is seen face-on and we have a circularly-polarized wave: $\langle h_+ \rangle = \langle h_\times\rangle$ where $\langle \ldots \rangle$
denote time-averages over one orbit. Note that as $f$ increases in time, so does $h(t)$ hence the characteristic ``chirping'' of GW signals.

$\Phi_\text{N}(t)$ in Eqs.~(\ref{eq:hplus_TD}, \ref{eq:hcross_TD}) is the phase of the GWs given by
%
\begin{align}
\Phi_\text{N}(t) &=\int_{t_i}^t dt' \omega(t') \nn\\
&= -2\left(\f{5 G M_c}{c^3}\right)^{-5/8} (t_\text{coal}-t)^{5/8}+ \Phi_0 \, ,\label{eq:Phase}
\end{align}
%
where $\Phi_0 \equiv \Phi(t=t_\text{coal})$ is an integration constant. The subscript N denotes the Newtonian (leading-order) contribution. 
Higher-order contributions can be added in terms a post-Newtonian series, which make up to $\lesssim 2\%$ of the total phase.
Although we will include contributions up to and including 3.5PN to obtain the results of Sec.~\ref{sec:results}, 
we will not show these expressions here, but they can be found in Ref.~\cite{Blanchet_LRR}.

@@

It remains to compute $|Q(\theta,\phi,\psi,\iota)|^2$. 
Since we can not a priori know the direction of the source, it is appropriate to average over the angles $\theta,\phi,\psi,\iota$ which we compute
%For our computations, we will only work with one detector so we can set $\psi=0$ without loss of generality, which results in
%%
%
%\be
%F_+(\theta,\phi) = \f{1}{2}(1+\cos^2\theta)\cos2\phi, \qquad F_\times(\theta,\phi) = \cos\theta \sin2\phi.
%\ee
%
%%
via the following RMS-averaging integral
%%
%
\be
\langle F^2_{+}\rangle \equiv \f{1}{2\pi}\int_0^{2\pi}d\psi \,\f{1}{4\pi}\int_0^{2\pi} d\phi\,\int_0^\pi d\theta \sin\theta\ F^2_{+}(\theta,\phi,\psi)=\f{1}{5}\label{eq:ang_av}
\ee
%
%%
and likewise $\langle F^2_{\times}\rangle=1/5 $.
Finally, we average over the orbital inclination angle
%
\be
%\langle g(\iota) \rangle = 
\f{1}{2} \int_{-1}^1 d(\cos\iota)\left[\left(\f{1+\cos^2\iota}{2}\right)^2+\cos^2\iota\right] = \f{4}{5} .\label{eq:inc_ang_av}
\ee
Thus we have that
%
\be
\langle |Q(\theta,\phi,\psi,\iota)|^2 \rangle_{\theta,\phi,\psi,\iota} = \f{4}{25}\label{eq:RMS_th_ph_ps_iota}.
\ee
%
As angles have zero averages in general, we can think of the RMS as the standard deviation ($\sigma =2/5$). 
%Therefore, $2\sigma= 4/5$ above the average (zero in this case) should statistically cover 95\% of the four-dimensional angle space. 
%Finally, we add the case of an optimally-oriented, purely $+$-polarized source ($\psi=0,\iota=0$) that is directly overhead ($\theta=0$) simply resulting in 
%
%\be
%\langle |Q(\theta=0,\phi=0,\psi=0,\iota=0)|^2 \rangle = 1.
%\ee
%
%
%
%\be
%\tilde{H}(f)\equiv  A\,\left\langle\, |Q(\theta,\phi,\psi,\iota)|^2\,\right\rangle^{1/2}\, \f{c}{D}\left(\f{G M_c}{c^3}\right)^{5/6} f^{-7/6}\, . \label{eq:ang_av_strain_FD}
%\ee
%
%%

@@

\section{Simplifications and idealizations used in evolving the inspiral}\label{sec:idealizations}

@@

Including the error bars, the known range for neutron star masses runs from $\approx 0.5 M_\odot$ to $ \approx 3 M_\odot$ with the actual values restricted
between $\approx 1 M_\odot$ and $\approx 2 M_\odot$ \cite{Ozel_Freire} (one exception being J1748-2021B with mass $2.74^{+0.21}_{-0.21} M_\odot $ \cite{Freire_et_al_2007}).

@@

Therefore, let us restrict the masses to lie between $1.2 M_\odot$ and $1.6 M_\odot$ then focus on % ranges $R_{1,2} \in [1,2]/(1.4)$. 

@@

$m_1=m_2=1.4 M_\odot$ is more likely to underestimate $T_\text{AW}$ than overestimate it which is something we can live with.

@@

>>>>

\begin{abstract}
We explore the intriguing possibility of employing future ground-based gravitational-wave interferometers to detect the inspiral of binary neutron stars sufficiently
early to alert electromagnetic observatories so that a gamma-ray burst (GRB) can be observed in its entirety from its very beginning.
We quantify the ability to predict a GRB by computing the time a binary neutron star (BNS) system takes to inspiral from its moment of detection to its final merger. We define the moment of detection to be the instant at which the inteferometer network accumulates a signal-to-noise ratio of 15. %for the BNS inspiral.
For our computations, %of advance warning times, 
we specifically consider BNS systems at luminosity distances of (i) $D\le200\,$Mpc in the three-interferometer Advanced-LIGO-Virgo network of 2020, and (ii) $D \le 1000\,$Mpc in the Einstein Telescope's B and C configurations. 
In the case of Advanced LIGO-Virgo we find that we may at best get a few minutes of warning time, thus we expect no forecast of GRBs in the 2020s. 
On the other hand, Einstein Telescope will provide us with advance warning times of up to five hours.  
Taking one hour as a benchmark warning time, we obtain a corresponding horizon
distance of roughly 600 Mpc for the Einstein Telescope C configuration.
%, more specifically, greater than an hour
%for BNS inspirals closer than roughly 600\,Mpc. Taking this luminosity distance to be our horizon distance, 
Using current BNS event rates within this horizon distance, we show that this translates to forecasts of $\gtrsim \ord(10^2)$ GRBs in the 2030s. %with Einstein C. %with the C configuration of Einstein Telescope.
We reapply our warning-time computation to binary black hole - neutron star inspirals and find that we expect 1 to 3 tidal disruption events 
to be forecast by the same detector.
%Our main conclusions are: (i) we expect ALV network to forecast 1 to 3 GRBs in the 2020s, (ii)

@@

GW170817 was initially identified by the LIGO-Hanford interferometer (H1) 
using a template bank of gravitational waveforms computed within the framework of post-Newtonian theory \cite{Buonanno:2009zt, Blanchet_LRR}.
The inspiral swept across the Advanced Ligo-Virgo (ALV) network's bandwidth from approximately $ 30\,$Hz to about $ 2\,$kHz in 57 seconds,
executing nearly $ 3000$ gravitational wave cycles and accumulating a network signal-to-noise ratio of 32.4 with a false-alarm rate of one per $8.0\times 10^4$ years \cite{GW170817}. 
The total mass of the system was inferred to be $2.74^{+0.04}_{-0.01} M_\odot$ with component masses in the range $1.17-1.60 M_\odot$, consistent
with neutron stars (assuming low spin priors, see Sec.~\ref{sec:idealizations}).
Moreover, both the gravitational wave (GW) and EM observations were consistent with a source luminosity distance of $\sim\!\! 40\,$Mpc
and supported the hypothesis that GW170817 resulted from the inspiral and merger of two neutron stars, 
causing the prompt short-hard gamma-ray burst (GRB170817A) in NGC4993 as first suggested by Ref.~\cite{Eichler:1989ve}.

@@

Given the sensitivities of interferometers to date, efforts have thus far mostly focused on reducing the latency of EM follow-ups, 
i.e., the lag time between the merger and the alerting of the EM observatories. 
LIGO-Virgo science runs in 2009-2010 yielded a latency of $\sim\!\! 30$ to $60\,$minutes \cite{Abbott:2011ys} (no detections).
In the case of GW170817, the latency was 40\,minutes and $36^{+1.7}\,$seconds \cite{GBM:2017lvd}.
It then took an additional $ 10$ hours to locate the optical transient. %designated AT 2017gfo.

@@

This article is organized as follows. Sec.~\ref{sec:BNS_inspiral} introduces the formulation for the leading-order evolution of the binary inspiral.
Sec.~\ref{sec:IFO_response} details the interferometer response to GWs based on topology.
Sec.~\ref{sec:idealizations} lists the various idealizations we employ to simplify our treatment and supplies justification for each.
Sec.~\ref{sec:GW170817} tests our model network and evolution using the parameters of GW170817 and the corresponding observations.
Sec.~\ref{sec:results} contains our main results presented as advance warning times in Tables~\ref{table:LIGO2020} and \ref{table:ET}; 
and as horizon distances and event rates in Table~\ref{table:horizon}.
In Sec.~\ref{sec:BH_NS}, we recompute the advance warning times in the case of black hole - neutron star inspirals to see whether or not 
future detectors can forecast a tidal disruption event, for which we summarize our findings in Table~\ref{table:ET_BH_NS}.

@@

\section{Binary neutron star inspirals}\label{sec:BNS_inspiral}
\subsection{Newtonian evolution}
For us, the starting point for the evolution of a binary under gravitational radiation reaction
%the emission of GWs 
is the expression for the power emitted in GWs given by 
the celebrated Einstein quadrupole formula \cite{1918SPAW}
%for the generated by the time-variance of the mass quadrupole moment
%
\be
\dot{E} = \f{G}{5c^5}\langle \dddot{Q}_{ij}\dddot{Q}_{ij} \rangle \label{eq:Edot_quadrupole}\, .
\ee
%
Here, $Q_{ij}$ is a trace-reversed {\bf MAYBE: explain this} mass quadrupole moment which
is the dominant contribution to the power emitted %in gravitational radiation
by a system with changing mass moments. Details of how to obtain this expression can be found in Ref.~\cite{Maggiore}.

Here, we are specifically interested in the motion of two point masses in a circular orbit around a common center of mass.
Labelling the masses by $m_1 \le m_2$, we have, for a binary in a circular orbit with Keplerian angular frequency $\Omega$ and separation $r$,
%
\be
\dot{E} = \f{32}{5}\f{G\mu^2}{c^5}\, r^4 \Omega^6\label{eq:Edot},
\ee
%
where $\mu= m_1 m_2/M$ is the reduced mass and $M=m_1+m_2$ is the total mass. 
We justify our simplification of the motion to circular orbits in Sec.~\ref{sec:idealizations}.
%Our assumption of the circularity of the motion is well justified in

An important quantity in GW astronomy is the chirp mass of the binary given by
%
\be
M_c = \mu^{3/5} M^{2/5} = \f{(m_1 m_2)^{3/5}}{(m_1+m_2)^{1/5}} \label{eq:chirp_mass}\, ,
\ee
%
which turns Eq.~(\ref{eq:Edot}) into
%
\be
\dot{E} = \f{32}{5}\f{c^5}{G} \left(\f{G M_c\, \omega}{2c^3}\right)^{10/3}, \label{eq:Edot2}
\ee
%
where $\omega=2\Omega$ is the angular frequency of the quadrupolar GWs. 
The quantity in parantheses in Eq.~(\ref{eq:Edot2}) equals $x^{3/2}$ where 
$x \equiv ({G M_c\, \Omega}/{c^3})^{2/3} $
is the dimensionless inverse separation parameter ubiquitous in post-Newtonian theory.
As $x\propto v^2/c^2$, it is used to keep track of post-Newtonian orders and we can see from Eq.~(\ref{eq:Edot2}) that $\dot{E}\propto x^5$. 
%hence the common statement in the literature saying that gravitational radiation starts

@@

Let us now turn our attention to the GWs generated by the inspiral.
These are tensorial perturbations propagating at the speed of light in flat (Minkowski) spacetime.
In the so-called wave zone, defined by the condition $ \lambda= c/f \ll D$, where $\lambda$ is the GW wavelength and $D$ is the distance to the GW source, 
these tensor modes satisfy the flat spacetime sourceless wave equation. %$\Box\bar{h}_{\mu\nu} = 0$.
%meaning that we are considering the solutions far from the GW source, 
In a suitable gauge, such as the commonly used transverse-traceless gauge, 
it can be shown that there exist only two physical degrees of freedom which manifest themselves as two independent polarization amplitudes $h_+$ and $h_\times$. %when interacting with masses away from the source.
For a binary in a circular orbit, these states read
%
\begin{align}
 h_+(t) &= h_c(t)\, \left(\f{1+\cos^2\iota}{2}\right)\, \cos[\Phi_\text{N}(t)],\label{eq:hplus_TD}\\
 h_\times(t) & =h_c(t)\,\cos\iota \sin[\Phi_\text{N}(t)]\label{eq:hcross_TD},
\end{align}
%
%
where
%
\be
h_c(t) =\f{4}{D}\left(\f{G M_c}{c^2}\right)^{5/3}\, \left(\f{\pi f(t)}{c}\right)^{2/3} \label{eq:h_c}
\ee
is the characteristic strain and
$\iota=\cos^{-1}(\mathbf{\hat{n}}\cdot{\mathbf{\hat{L}}}) $ is the inclination angle between the line of sight unit vector $\mathbf{\hat{n}} $ and the orbital angular momentum unit vector $\mathbf{\hat{L}}$. 
For an orbit seen edge on, $\iota = \pi/2$ giving $h_\times=0$, i.e,  a purely plus-polarized wave.
When $\iota=0$ the orbit is seen face-on and we have a circularly-polarized wave: $\langle h_+ \rangle = \langle h_\times\rangle$ where $\langle \ldots \rangle$
denote time-averages over one orbit. Note that as $f$ increases in time, so does $h(t)$, hence the characteristic ``chirping'' of GW signals.

$\Phi_\text{N}(t)$ in Eqs.~(\ref{eq:hplus_TD}, \ref{eq:hcross_TD}) is the phase of the GWs given by
%
\begin{align}
\Phi_\text{N}(t) &=\int_{t_i}^t dt' \omega(t') \nn\\
&= -2\left(\f{5 G M_c}{c^3}\right)^{-5/8} (t_\text{coal}-t)^{5/8}+ \Phi_0 \, ,\label{eq:Phase}
\end{align}
%
where $\Phi_0 \equiv \Phi(t=t_\text{coal})$ is an integration constant. The subscript N denotes the Newtonian (leading-order) contribution. 
Higher-order contributions can be added in terms of a post-Newtonian series, which make up to $\lesssim 2\%$ of the total phase.
Although we will include contributions up to and including 3.5PN to obtain the results of Sec.~\ref{sec:results}, 
we will not show these expressions here, but they can be found in Ref.~\cite{Blanchet_LRR}.

@@

It remains to compute $|Q(\theta,\phi,\psi,\iota)|^2$. 
Since we can not a priori know the direction of the source, it is appropriate to average over the angles $\theta,\phi,\psi$, and $\iota$ which we compute
%For our computations, we will only work with one detector so we can set $\psi=0$ without loss of generality, which results in
%%
%
%\be
%F_+(\theta,\phi) = \f{1}{2}(1+\cos^2\theta)\cos2\phi, \qquad F_\times(\theta,\phi) = \cos\theta \sin2\phi.
%\ee
%
%%
via the following RMS-averaging integral
%%
%
\be
\langle F^2_{+}\rangle \equiv \f{1}{2\pi}\int_0^{2\pi}d\psi \,\f{1}{4\pi}\int_0^{2\pi} d\phi\,\int_0^\pi d\theta \sin\theta\ F^2_{+}(\theta,\phi,\psi)=\f{1}{5}\label{eq:ang_av}
\ee
%
%%
and likewise $\langle F^2_{\times}\rangle=1/5 $.
Finally, we average over the orbital inclination angle
%
\be
%\langle g(\iota) \rangle = 
\f{1}{2} \int_{-1}^1 d(\cos\iota)\left[\left(\f{1+\cos^2\iota}{2}\right)^2+\cos^2\iota\right] = \f{4}{5} .\label{eq:inc_ang_av}
\ee
Thus we have that
%
\be
\langle |Q(\theta,\phi,\psi,\iota)|^2 \rangle_{\theta,\phi,\psi,\iota} = \f{4}{25}\label{eq:RMS_th_ph_ps_iota}.
\ee
%
As angles have zero averages in general, we can think of the RMS as the standard deviation ($\sigma =2/5$). 
%Therefore, $2\sigma= 4/5$ above the average (zero in this case) should statistically cover 95\% of the four-dimensional angle space. 
%Finally, we add the case of an optimally-oriented, purely $+$-polarized source ($\psi=0,\iota=0$) that is directly overhead ($\theta=0$) simply resulting in 
%
%\be
%\langle |Q(\theta=0,\phi=0,\psi=0,\iota=0)|^2 \rangle = 1.
%\ee
%
%
%
%\be
%\tilde{H}(f)\equiv  A\,\left\langle\, |Q(\theta,\phi,\psi,\iota)|^2\,\right\rangle^{1/2}\, \f{c}{D}\left(\f{G M_c}{c^3}\right)^{5/6} f^{-7/6}\, . \label{eq:ang_av_strain_FD}
%\ee
%
%%

@@

\section{Simplifications and idealizations} %used in evolving the inspiral}
\label{sec:idealizations}

@@

Including the error bars, the known range for neutron star masses runs from $\approx 0.5 M_\odot$ to $ \approx 3 M_\odot$ with typical values between $\approx 1 M_\odot$ and $\approx 2 M_\odot$ \cite{Ozel_Freire} (one exception being J1748-2021B with mass $2.74^{+0.21}_{-0.21} M_\odot $ \cite{Freire_et_al_2007}).

@@

Therefore, let us restrict the masses to lie between $1.2 M_\odot$ and $1.6 M_\odot$, then focus on % ranges $R_{1,2} \in [1,2]/(1.4)$. 

@@

$m_1=m_2=1.4 M_\odot$ is more likely to underestimate $T_\text{AW}$ than overestimate it. 
%which is something we can live with.

@@

Let us briefly discuss the possibility of eccentric BNS systems {\bf FINISH}
\cite{1975MNRAS.172P..15F,Grindlay:2005ym}
